---
title: "Digit Recognizer"
output: html_document
smart : FALSE
---

This is from Kaggle Titanic Contest. I am trying to analyze the data and answer the questions.

```{r}
#Read the dataset
library(caret)
library(kernlab)
library(stringr)

digits = read.csv("train.csv")

set.seed(1680) # for reproducibility

library(dplyr) # for data cleaning
library(ISLR) # for college dataset
library(cluster) # for gower similarity and pam
library(Rtsne) # for t-SNE plot
library(ggplot2) # for visualization

table(as.factor(digits$label))

ggplot(digits, aes(x = as.factor(label), fill = label)) +
  geom_bar(stat = "count", color = "white") +
  scale_fill_gradient(low = "lightblue", high = "pink", guide = FALSE) +
  labs(title = "Digits in Train Data", x = "Digits")

```


```{r}

sample <- sample(1:nrow(digits), 50)
var <- t(digits[sample, -1])
var_matrix <- lapply(1:50, function(x) matrix(var[, x], ncol = 28))
opar <- par(no.readonly = T)
par(mfrow = c(5, 10), mar = c(.1, .1, .1, .1))

for (i in 1:50) {
    for (j in 1:28) {
        var_matrix[[i]][j,] <- rev(var_matrix[[i]][j,])
    }
    image(var_matrix[[i]], col = grey.colors(225), axes = F)
}

```

Find and remove columns with near Zero Variance

```{r}
nzr <- nearZeroVar(digits[, -1], saveMetrics = T, freqCut = 10000 / 1, uniqueCut = 1 / 7)
sum(nzr$nearZeroVar)
```

Remove the near zero columns
```{r}
cutvar <- rownames(nzr[nzr$nzv == TRUE,])
var <- setdiff(names(digits), cutvar)
digits <- digits[, var]
```

Get Covariance Train
```{r}
label <- as.factor(train[[1]])
digits$label <- NULL
digits <- digits / 255
covtrain <- cov(digits)
```

Do PCA
```{r}

train_pc <- prcomp(covtrain)
varex <- train_pc$sdev ^ 2 / sum(train_pc$sdev ^ 2)
varcum <- cumsum(varex)
result <- data.frame(num = 1:length(train_pc$sdev),
                         ex = varex,
                         cum = varcum)


ggplot(result, aes(y = cum, x = num)) + geom_point() + geom_vline(xintercept = 25);

#plot(result$num, result$cum, type = "b", xlim = c(0, 100),
     #main = "Variance Explained by Top 100 Components",
     #xlab = "Number of Components", ylab = "Variance Explained")
#abline(v = 25, lty = 2)
```

Visualize PCA output
```{r}
train_score <- as.matrix(digits) %*% train_pc$rotation[, 1:25]
digits <- cbind(label, as.data.frame(train_score))

colors <- rainbow(length(unique(digits$label)))
names(colors) <- unique(train$label)
ggplot(digits, aes(x = PC2, y=PC3)) + geom_point(aes(color = label))

#text(train$PC1, train$PC2, label = train$label, col = colors[train$label])
```


Train and predict
```{r}
svm_mdl <- train(label ~ ., data = digits,
                 method = "svmRadial",
                 trControl = trainControl(method = "cv",
                                        number = 5),
                 tuneGrid = data.frame(sigma = 0.01104614,
                                      C = 3.5))
svm_mdl
```


Predict the digits
```{r}
test = read.csv("test.csv")
test <- test[, var[-1]] / 255
test <- as.matrix(test) %*% train_pc$rotation[, 1:25]
test <- as.data.frame(test)

pred <- predict(svm_mdl$finalModel, test, type = "response")
prediction <- data.frame(ImageId = 1:nrow(test), Label = pred)

write.csv(prediction, file = "DigitsPrediction_12_3.csv", row.names = FALSE)
```

* Analyze the data 
* Compute/Impute missing values
* Analyze Predictors
* Predict and Report

Do image processing to identify digits

